Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-06-27 sattose/paper.tex
Merge -> Refactor

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolpceurws}
\usepackage{float}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\newfloat{eqfloat}{h}{eqflts}
\floatname{eqfloat}{Equation}

\title{Towards Automated Merging of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -}

@@

\section{Addressing problems with clone type definitions}\label{chap:clonetypes}
For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings (see section \ref{chap:backgroundclonetypes} for these definitions) to increase the chance that we can merge the clone while improving the design. Due to the serious challenges involved in their detection and refactoring, type 4 clones are not considered in this study.

@@

We tackle these problems with type 2R clones to be able to detect such clones that can and should be merged. Our definition ensures functional similarity by applying the following changes to type 2 clones:

@@

\subsection{Relations Between Clone Instances} \label{chap:relationsinstances}
When merging code clones in object-oriented languages, it is very important to consider the relation between clone instances. This relation has a big impact on how a clone should be merged, in order to improve the software design in the process. In this section, we display measurements we conducted on the corpus introduced in section \ref{chap:corpus}. These measurements are based on an experiment by Fontana et al. \cite{fontana2015duplicated}, which we will briefly introduce in section \ref{chap:catcloneinstancerelations}. We use a vastly different setup, which is explained in section \ref{chap:oursetup}. We then show our results in section \ref{chap:ourmeasurements}.

@@

\section{Merging duplicate code through method extraction}\label{chap:mergingdups}
The most used technique to merge clones is method extraction (creating a new method on basis of the contents of clones). However, method extraction cannot be applied in all cases. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, the cloned code can contain statements like \texttt{return}, \texttt{break}, \texttt{continue}. In these instances, more conditions may apply to be able to conduct a refactoring, if beneficial at all.

We measured the amount of clones that can be refactored through method extraction (without additional transformations being required). Our results are displayed in Table \ref{table:refactorability}. In this table we use the following categories:
\begin{itemize}
    \item \textbf{Can be extracted:} This clone is a fragment of code that can directly be extracted to a method. Then, based on the relation between the clone instances, further refactoring techniques can be used to merge the extracted methods (for instance ``pull up method'' for clones in sibling classes).
    \item \textbf{Complex control flow:} This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements.
    \item \textbf{Spans part of a block:} This clone spans a part of a statement.
    \item \textbf{Is not a partial method:} If the clone does not fall in the ``Partial method'' category of Table \ref{table:contents}, the ``extract method'' refactoring technique cannot be applied.
\end{itemize}

@@

We have looked into existing definitions for different types of clones \cite{roy2007survey} and proposed solutions for problems that these types have with regards to automated refactoring. We propose that fully qualified identifiers of method call signatures and type references should be considered instead of their plain text representation, to ensure refactorability. Furthermore, we propose that one should define thresholds for variability in variables, literals and method calls, in order to limit the number of parameters that the merged unit shall have.

>>>>

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolpceurws}
\usepackage{float}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{tikz}
\newfloat{eqfloat}{h}{eqflts}
\floatname{eqfloat}{Equation}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

\title{Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -}

@@

\section{Addressing problems with clone type definitions}\label{chap:clonetypes}
For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings (see section \ref{chap:backgroundclonetypes} for these definitions) to increase the chance that we can refactor the clone while improving the design. Due to the serious challenges involved in their detection and refactoring, type 4 clones are not considered in this study.

@@

We tackle these problems with type 2R clones to be able to detect such clones that can and should be refactored. Our definition ensures functional similarity by applying the following changes to type 2 clones:

@@

\subsection{Relations Between Clone Instances} \label{chap:relationsinstances}
When merging code clones in object-oriented languages, it is very important to consider the relation between clone instances. This relation has a big impact on how a clone should be refactored, in order to improve the software design in the process. In this section, we display measurements we conducted on the corpus introduced in section \ref{chap:corpus}. These measurements are based on an experiment by Fontana et al. \cite{fontana2015duplicated}, which we will briefly introduce in section \ref{chap:catcloneinstancerelations}. We use a vastly different setup, which is explained in section \ref{chap:oursetup}. We then show our results in section \ref{chap:ourmeasurements}.

@@

\section{Merging duplicate code through method extraction}\label{chap:mergingdups}
The most used technique to refactor clones is method extraction (creating a new method on basis of the contents of clones). However, method extraction cannot be applied in all cases. Sometimes a clone spans a statement partially (like a for-loop of which only it's declaration and a part of the body is cloned). Merging the clones can be harder in such instances. Also, the cloned code can contain statements like \texttt{return}, \texttt{break}, \texttt{continue}. In these instances, more conditions may apply to be able to conduct a refactoring, if beneficial at all.

We measured the amount of clones that can be refactored through method extraction (without additional transformations being required). Our results are displayed in Table \ref{table:refactorability}. In this table we use the following categories:
\begin{itemize}
    \item \textbf{Can be extracted:} This clone is a fragment of code that can directly be extracted to a method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance ``pull up method'' for clones in sibling classes).
    \item \textbf{Complex control flow:} This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements.
    \item \textbf{Spans part of a block:} This clone spans a part of a statement.
    \item \textbf{Is not a partial method:} If the clone does not fall in the ``Partial method'' category of Table \ref{table:contents}, the ``extract method'' refactoring technique cannot be applied.
\end{itemize}

@@

We have looked into existing definitions for different types of clones \cite{roy2007survey} and proposed solutions for problems that these types have with regards to automated refactoring. We propose that fully qualified identifiers of method call signatures and type references should be considered instead of their plain text representation, to ensure refactorability. Furthermore, we propose that one should define thresholds for variability in variables, literals and method calls, in order to limit the number of parameters that the refactored unit shall have.
